---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# constructive

<!-- badges: start -->
<!-- badges: end -->

`dput()` but better

## Examples

```{r, error = TRUE}
library(constructive)

foo <- data.frame(
  a = factor(month.abb),
  b = as.Date(sprintf("2022-%s-01", 1:12))
)

dput(foo)

construct(foo)

# we don't have to deparse everything
construct(foo, data = list(month.abb = month.abb)) # or dplyr::lst(month.abb)

# we'll often use `dplyr::lst` and we can even use some complex expressions there
construct(list(c("a", "b", "c"), c("d", "e", "f")), data = dplyr::lst(head(letters, 3)))

# we can handle some complex attributes
library(dplyr, warn = F)
grouped_band_members <- group_by(band_members, band)
construct(grouped_band_members)

# And we can implement alternative constructors
construct(grouped_band_members, tribble = TRUE)

# the data arg can be a package name too
construct(grouped_band_members, data = "dplyr")

# dms are supported too
library(dm, warn = F)
construct(dm(cars = head(cars)))

construct(dm_pixarfilms(), data = "pixarfilms")

# environments are not always possible to reproduce but we support some common cases
search()
construct(as.environment(search()[1]))
construct(as.environment(search()[2]))
construct(as.environment(search()[12]))
construct(as.environment(search()[13]))
construct(environment(group_by))

# and if not possible, we try something that might sometimes be enough
e <- new.env()
e$x <- 1
e$y <- 2
construct(e)

# We can construct some functions faithfully, using `rlang::new_function()`
construct(group_by)
```

# How it works

{constructive} has a main exported function `construct()` built around a `construct_raw()`, the main
unexported function, and adding some checks and pretty printing using {styler}.

`construct_raw` is called recursively through list-like objects and is made of 3 steps :

* Check if we already have the object in store in our `data` arg, if so display its name rather than the code to rebuild it
* If the object cannot be found in `data` build it idiomatically by calling the `construct_idiomatic()`
 generic
* Then the `repair_attributes()` generic adapts the above to make sure created object
has the same class and other attributes as the source object
 
To extend the package to a new object we only need to add a method for `construct_idiomatic()`
and for `repair_attributes()`.

# Use cases 

* create reproducible examples
* compact snapshot tests that contain all the information
* debugging (avoiding the caveats of the default printing, where 2 objects might print the same but not be the same)
* Understand object structure better than when using `dput` or `str`, even if
  {constructive} doesn't feature a specific constructor for the class.

```{r}
x <- table(c(1, 1, 1, 4))

x

construct(x)

# compare with `str`, a generic which doesn't display everything:
str(x)

# and `dput()` which is often confusing, and sometimes uses different attribute names
dput(x)

# we can use `max_atomic` to hide all atomic vectors and see only the skeleton:
construct(table(c(1,1,1,4)), max_atomic = 0)

construct(iris, max_atomic = 2)
```

## construct_diff

An alternative to `waldo::compare()` (looks better in the IDE without `interactive = FALSE`) 

```{r}
# The args max_atomic, max_boy and env_as_list can be used to reduce output,
# here we want to see what adding geom_point() does to a ggplot
library(ggplot2)
construct_diff(
  # max_atomic limits number of displayed elements for atomic vectors
  max_atomic = 4, 
  # if env_as_list is FALSE instead of defining unnamed envs as 
  # `as.environment(list(elt = ...))` we simply use `new.env()`
  env_as_list = FALSE, 
  interactive = FALSE,
  ggplot(cars, aes(speed, dist)),
  ggplot(cars, aes(speed, dist)) + geom_point()
)
```

