---
title: "Extend constructive"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extending constructive}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(constructive)
```

We detail in this vignette how {constructive} works and how
you might extend it by defining custom  `.cstr_construct.<class>()` and 
`.cstr_construct.<class>.<constructor>()` methods.

The functions `.cstr_new_class()` and `.cstr_new_constructor()` can be used to
create custom constructors from templates. The easiest workflow is probably
to take a look at the 
package [{constructive.examples}](https://github.com/cynkra/constructive.example/) which will guide you through the process, then call these functions with the argument `commented = TRUE`.

If this is not enough, or if you want to know more, we describe below in 
more details how the package and its key functions work.

## `construct()` and `.cstr_construct()`

* `.cstr_construct()` builds code recursively, without checking the inputs or output 
  validity, without error handling, and without formatting.
* `construct()` wraps `.cstr_construct()` and does this extra work.
* `.cstr_construct()` is a generic and many methods are implemented in the package, for
  instance `construct(iris)` will call `.cstr_construct.data.frame()` which down the line
  will call `.cstr_construct.numeric()` and `.cstr_construct.factor()` to construct its columns.
* `.cstr_construct()` contains extra logic to :
  * Attempt to match its data input to a list of objects provided to the `data` argument.
  * Restrict the dispatch so the `classes` argument and the functions `construct_base()` and
    `construct_dput()` can work.
  
```{r}
.cstr_construct
# a character vector
.cstr_construct(letters)
# a constructive object, 
construct(letters)
```

## `.cstr_construct.<class>()` methods

`.cstr_construct.<class>()` methods are generics themselves, they typically have this form:

```{r}
.cstr_construct.Date <- function(x, ...) {
  opts <- list(...)$opts$Date %||% opts_Date()
  if (is_corrupted_Date(x) || opts$constructor == "next") return(NextMethod())
  UseMethod(".cstr_construct.Date", structure(NA, class = opts$constructor))
}
```

* `opts` contains options passed to the `...` or `template` arguments of 
  `construct()` through the `opts_*()` functions, we fall back to a default value if none were provided.
  The `list(...)$opts` idiom is used in various places, it allows us to forward
  the dots conveniently. 
* If the object is corrupted or if the user decided to bypass the current method by
  choosing "next" as a constructor for this class, we return `NextMethod()` to
  forward all our inputs to a lower level constructor.
* We finally dispatch to a method based on the constructor, and not on the object's
  class as is done traditionally. This explains the unusual look of the `UseMethod()`
  call above.
  
## `opts_?()` function

When implementing a new method you'll need to define and export the corresponding
`opts_?()` function. It provides to the user with a way to choose a constructor and 
additional parameters, and sets the default behavior.

It should always have this form:

```{r, eval=FALSE}
opts_Date <- function(
    constructor = c(
      "as.Date", "as_date", "date", "new_date", "as.Date.numeric", "as_date.numeric", "next", "double"), 
    ..., 
    origin = "1970-01-01") {
  .cstr_options("Date", constructor = constructor[[1]], ..., origin = origin)
}
```

* The class is present in the name of the `opts_?()` function and as the first
  argument of `.cstr_options()`.
* A character vector of constructors is provided, starting with the default constructor
* A "next" value is mandatory except for internal types.
* Additional options retrievable by the constructor might be added, as we did here with
  `origin`
* Note that we don't use `match.arg()` here, because new constructors can be defined
  outside of the package.

## `is_corrupted_?()` function

`is_corrupted_?()` checks if `x` has the right internal type and attributes,
sometimes structure, so that it satisfies the expectations of a well formatted object
of a given class.

If an object is corrupted for a given class we cannot use constructors for this
class, so we move on to a lower level constructor by calling `NextMethod()` in
`.cstr_construct()`.

This is important so that `{constructive}` doesn't choke on corrupted objects but
instead helps us understand them.

For instance in the following example `x` prints like a date but it's corrupted,
a date should not be built on top of characters and this object cannot be built
with `as.Date()` or other idiomatic date constructors.

```{r, error = TRUE}
x <- structure("12345", class = "Date")
x
x + 1
```

We have defined : 

```{r}
is_corrupted_Date <- function(x) {
  !is.double(x)
}
```

And as a consequence the next method, `.cstr_construct.default()` will be called
through `NextMethod()` and will handle the object using an atomic vector constructor:

```{r}
construct(x)
```

## constructors

constructors are functions named as `.cstr_construct.<class>.<constructor>`.

For instance the default constructor for "Date" is :

```{r}
constructive:::.cstr_construct.Date.as.Date
```

Their arguments are `x` and `...`, and not more. Additional parameters fed
to the `opt_<class>()` function can be fetched from `list(...)$opts$<class>`

The function `.cstr_apply()` is used to construct arguments recursively.

Sometimes a constructor cannot handle all cases and we need to fall back to
another constructor, it happens above because `Inf`, `NA`, or decimal dates
cannot be represented by a string wrapped by `as.Date()`.

```{r}
x <- structure(c(12345, 20000), class = "Date")
y <- structure(c(12345, Inf), class = "Date")
construct(x)
construct(y)
```

That last line of the function is essential, it does the attribute repair.

## Attribute repair

Constructors should always end by a call to `.cstr_repair_attributes()` or a function
that wraps it.

These are needed to adjust the attributes of an object after idiomatic constructors such as `as.Date()` have defined their data and canonical attributes.

```{r}
x <- structure(c(12345, 20000), class = "Date", some_attr = 42)
# attributes are not visible due to "Date"'s printing method
x

construct(x)
```

`.cstr_repair_attributes()` essentially sets attributes with exceptions :

* It doesn't set names by default, these should be handled by the constructors
* It doesn't set the class explicitly if it's identical to the idiomatic class, i.e. the
  class returned by the constructor before the repair call, and provided through the
  `idiomatic_class` argument
* It doesn't set attributes that we choose to ignore because they are set by the
  constructor (e.g. row names for data frames or levels for factors)

```{r}
constructive:::repair_attributes_Date

constructive:::repair_attributes_factor

constructive:::repair_attributes_tbl_df
```

